Validating FF-Replan vs.\@ a State-Of-The-Art Probabilistic Planner in Translated Real World Domains

### 2014-03-30 FinalPaper/paper.tex
added initial part background

\section{Problem Formulation}

\subsection{Markov Decision Process}

\subsubsection{Goal-oriented MDPs}

\subsubsection{MDP with Finite Horizon}

\subsection{Planners}

\subsubsection{Classical Planners}

A planner tries to find a sequence of actions such that a goal or maximum
reward is achieved. A classical planning problem consists out of states,
actions, conditions and a goal state.

\subsubsection{Probabilistic Planning} Classical Planners expect a
deterministic world. For many Real World applications actions happen with a
certain probability. This kind of Probabilistic Planning Problems can be
described with Markov Decision Processes (MDPs). For the purpose of our
analysis we define the problem as:

\begin{itemize}
	\item finite set of states $S$
	\item an initial state $s_o \in S$
	\item a $G \subseteq S$ of (absorbing) goal states
	\item a set $O$ of outcomes; probability of $o \in O$ is $\Pr(o)$
	\item a (total) deterministic transition function $T(o,s) \in S$ for all
		outcomes $o \in O$ and states $s \in S$
	\item sets $A(s)$ of applicable actions for each $s \in S$, together with
		function $out(a) \subseteq O$ mapping each action to a set of outcomes
		such that each outcome $o \in O$ belongs exactly to one action $act(o)$
		and $\sum_{o \in out(a)} \Pr(o) = 1$ for all $a$
\end{itemize}

% depends if we should include this, what we tell about GOURMAND / RDDL...

Besides a single or multiple goals, some planning problems also include a
reward function: For every $s \in S$, $s' \in S$, and $a \in A(s)$,
$R(s,a,s')$ is the reward function. The goal if the problem is to maximize
the reward until a goal is reached. If a Finite Horizon (number of steps of the
plan) is used, the goal of the problem is to achieve the highest reward within
the horizon.

A probabilistic planner takes the probabilistic planning problem as input and
builds and outputs a plan for the problem.

In this paper we use GOURMAND as a representative probabilistic planner, as
it's the improved version of Glutton, the runner-up of the IPPC-11 competition.

\subsubsection{Replanning}

@@

\begin{itemize}
	\item Summarize shortcomings of the previous solutions
	\item Describe the planning problem
	\item Argument the choice for 2 current papers to analyze
	\item Define how we would like to test the planning algorithms in a domain by our choice
\end{itemize}

\subsection{Planning problem}

\begin{itemize}
	\item Describe the MDP finite horizon planning problem (useful for modelling nearly infinite processes)
	\item reason about the \emph{augmented state space} of M(H), which is a set \emph{S x (0, ... , H)} of state number of steps-to-go pairs
	\item Describe reward heuristic
	\item Solving M(H) means finding a policy: a rule for selecting actions in augmented states so that executing the actions recommended by the policy starting in the initial state will accumulate the largest expected reward over H steps time
	\item The challenge or large branching factors (induced by natural dynamic)
\end{itemize}

@@

\section{Planners}

>>>>

\section{Background}

\subsection{Planning problems}
\subsubsection{Classical Planning problems. } Planning is the process of computing several steps of a problem-solving procedure before executing any of them. In classical planning problems the environment is assumed to \emph{deterministic, observable and state}. Classical planning problems consists out of states, actions, conditions and a goal state and are solved using search and can be aided using heuristics. The planner tries to find a sequence of actions such that a goal or maximum reward is achieved.

\subsubsection{Probabilistic Planning problems} 
Real-World applications are far from deterministic, induced by natural dynamics and chance. To accomodate for Real-World domains probabilistic planning problems were introduced.  These systems can be described using \emph{Markov Decision Processes} (MDPs) and can account for observable uncertainty. Probabilistic planning problems can be solved by evaluating iterations and creating policies to reach a goal state, either online or offline.

\subsection{Problem domains}
\subsubsection{Goal-oriented MDPs. }The problems in IPPC-04 were specified using the Planning Domain Description Language (PPDLL). PPDDL specifies a probabistic planning problem as a set of probabilistic action sets and an initial and goal state description in which every action has a certain probability assigned to it \cite{FFReplan}. These MDPs are of the form $\mathit{M = (S, A, app, Pr, R)}$ where $\mathit{S}$ and $\mathit{A}$ are a finite set of states and actions, $\mathit{app(s)}$ is the set of all actions applicable in $\mathit{s}$. $\mathit{Pr}$ represents the probability of a state transaction $\mathit{(s, a, s')}$ for $\mathit{a}$ and $\mathit{R(s, a, s')}$ is the reward of applying $\mathit{a}$ to $\mathit{s}$ and generating $\mathit{s'}$ The planning problem is defined as a triple $\mathit{P = (M, s\textsubscript{0}, G, \rho)}$ where $\mathit{M = (S, A, app, Pr, R)}$ is an MDP, $\mathit{s\textsubscript{0} \in S}$ is the initial state, $\mathit{G \subseteq S}$ is the set of goal states and $\mathit{0 < \rho < 1}$ is a probability threshold \cite{teichteil2010incremental}.

 The planners from IPPC-04 were evaluated on how many trials reached the goal per problem and the average number of actions selected when succesfull. These planning problems were however relatively easy to solve as they did not account for \emph{natural dynamics} and/or large branching factors that occur in more realtistic systems \cite{kolobov2012glutton}. FF-Replan \cite{FFReplan}, as specified in Section \ref{sec:planners} kindly remembered the committee of the IPPC-04 that the specified planning problems that contained no dead-ends were fairly easy to solve using relatively simple tools to handle with uncertainty.

\subsubsection{MDPs with Finite Horizon. }IPPC-11 decided to respond by introducing more realistic MDPs and their evaluation criteria by using MDPs with finite horizons. The planning problems are defined as tuples of the form \\ $\mathit{M\textsubscript{s0}(H) = \langle \langle S, A, T, R, s\textsubscript{0} \rangle , H \rangle }$ where $\mathit{S}$ s a finite set of states, $\mathit{A}$ a finite set of actions, $\mathit{T}$ is the transition function, similar to $\mathit{Pr}$ in the goal oriented MDP. $\mathit{R}$ represents the transitioning reward, $\mathit{s\textsubscript{0}}$ and $\mathit{H}$ is the number of decision epochs after which the process stops. Solving $\mathit{M(H)}$ means finding a policy for executing recommended state from the initial state over $\mathit{H}$ decision epochs \cite{kolobov2012gourmand}.

Translating finite horizon MDPs to goal-oriented domains results ends up in all state $H$ steps away from $s\textsubscript{0}$ are goal states. The evaluation criteria in finite horizon MDPs is the \emph{expected reward} and not the probability of the policy reaching the goal states.

\subsection{Planning algorithms}
\subsubsection{Replanning}

@@

\subsubsection{Probabilstic planning}
In this paper we use Gourmand as a representative probabilistic planner, as
it's the improved version of Glutton, the runner-up of the IPPC-11 competition.

@@

\section{Problem formulation}
\begin{itemize}
	\item Describe the MDP finite horizon planning problem (useful for modelling nearly infinite processes)
	\item reason about the \emph{augmented state space} of M(H), which is a set \emph{S x (0, ... , H)} of state number of steps-to-go pairs
	\item Describe reward heuristic
	\item Solving M(H) means finding a policy: a rule for selecting actions in augmented states so that executing the actions recommended by the policy starting in the initial state will accumulate the largest expected reward over H steps time
	\item The challenge or large branching factors (induced by natural dynamic)
	\item Summarize shortcomings of the previous solutions
	\item Describe the planning problem
	\item Argument the choice for 2 current papers to analyze
	\item Define how we would like to test the planning algorithms in a domain by our choice
\end{itemize}

\section{Planners}
\label{sec:planners}
